A Dependently Typed Multi-Stage Calculus

### 2019-07-30 reply-for-review.txt

Dear Reviewers,
Thank you for your detailed and valuable comments.

We will incorporate the comments as much as possible to a final version of the paper is accepted.  We will focus below on a few important comments (by Reviewers 1 and 2) we think we should address here.

------------------ Reviewer 1 -------------------------

Comment on the last sentence of the abstract:
This sentence refers to the "vadd" example and handling of the symbol of CSP in the type equivalence relation. We think they are an important part of this paper. We'll try to improve but probably will not remove it.

Comment on the last sentence before Section 3.2:
This statement is a conjecture but we think it is not difficult to prove.  We'll make it clear.

Comment on design choice to make type-level CSP:
Because we don't have staged semantics for types, explicit type-level CSP just makes the system more complicated.

Comment on the proof of Strong Normalisation:
Yes, SN for a calculus with pattern-matching can be proved as suggested.  We are not sure how the progress property contributes to the discussion here, though: what we would like to prove is a simulation property, which assumes one-step (successful) reduction in the source calculus and so the progress is not used.  (Are we missing anything?)

------------------ Reviewer 2 -------------------------

Comment on "A metaprogramming framework for formal verification" (by Ebner et al.) and "Ur: Statically-Typed Metaprogramming with Type-Level Record Computation" (by Chlipala):
The first paper is loosely related in the sense that they are interested in (automated) construction of well-typed terms but, in the MSP setting, the presence of eval (which it does not care) often complicates the problem a lot. The second paper is also loosely related because it focuses on practical use and does not contain a theory for metaprogramming.

Comment on "Program Logics for Homogeneous Generative Run-Time Meta-Programming"
(by Berger and Tratt): Although this paper takes a different approach from ours, its main purpose is similar to ours. The main difference is that their language is based on lambda-Box, which can manipulate only closed code, whereas ours can deal with open code.

### 2019-07-30 reply-for-review.txt

Thank you for your detailed and valuable comments.

We will incorporate all feedback and comments, including those not addressed
here. We will now respond to some of the key comments and questsions from each
reviewer.

@@

------------------ Reviewer 1 -------------------------

Comment on the last sentence of the abstract:
This sentence refers to the "vadd" example and handling of the symbol of CSP in
the type equivalence relation. We think they are an important part of this
paper. We'll try to improve but probably will not remove it.

Comment on the last sentence before Section 3.2:
This statement is a conjecture but we think it is not difficult to prove.
We will clarify this point.

Comment on design choice to make type-level CSP:
As we don't have staged semantics for types, explicit type-level CSP ultimately
complicates the system.

@@

Comment on the proof of Strong Normalisation:
Yes, SN for a calculus with pattern-matching can be proved as suggested.  We
are not sure how the progress property contributes to the discussion here,
though: what we would like to prove is a simulation property, which assumes
one-step (successful) reduction in the source calculus and so the progress is
not used.  (Are we missing anything?)

------------------ Reviewer 2 -------------------------

Comment on "A metaprogramming framework for formal verification" (by Ebner et
        al.) and "Ur: Statically-Typed Metaprogramming with Type-Level Record
        Computation" (by Chlipala):
The first paper is loosely related in the sense that they are interested in
(automated) construction of well-typed terms but, in the MSP setting, the
presence of eval (which it does not care) often complicates the problem a lot.
We will include this discussion in the Related work section.
The second paper is also loosely related because it focuses on practical use
and does not contain a theory for metaprogramming.

### 2019-07-30 reply-for-review.txt

Thank you for your detailed and valuable comments.

We will incorporate all feedback and comments, including those not addressed
here. We will now respond to some of the key comments and questions from each
reviewer.

@@

Comment on the proof of Strong Normalisation:
Yes, SN for a calculus with pattern-matching can be proved as suggested.  We
are not sure how the progress property contributes to the discussion here,
though: what we would like to prove is a simulation property, which assumes
one-step (successful) reduction in the source calculus and so the progress is
not used.  (Are we missing anything?)

------------------ Reviewer 2 -------------------------

Comment on "A metaprogramming framework for formal verification" (by Ebner et
        al.) and "Ur: Statically-Typed Metaprogramming with Type-Level Record
        Computation" (by Chlipala):
The first paper is loosely related in the sense that they are interested in
(automated) construction of well-typed terms but, in the MSP setting, the
presence of eval (which it does not care) often complicates the problem a lot.
We will include this discussion in the Related work section.
The second paper is also loosely related because it focuses on practical use
and does not contain a theory for metaprogramming.

Comment on the ability of this calculus as a model for a practical language
such as MetaOCaml:
Because it is easy to extend this calculus to include recursion, we think it is
able to use as a model of a practical language.

@@

Comment on "Program Logics for Homogeneous Generative Run-Time Meta-Programming"
    (by Berger and Tratt): 
Although this paper takes a different approach from ours, its main purpose is
similar to ours. The main difference is that their language is based on
lambda-Box, which can manipulate only closed code, whereas ours can deal
with open code.

------------------ Reviewer 3 -------------------------

Comment on the type checking algorithm:
We will construct the algorithm by extending an algorithm for dependent type
system.

### 2019-08-14 main.tex
Improving proof styles.
Add Acknowledgments.
small edit
Reflected comments on typing from reviwer 1,

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\newif\ifnotanonymous \notanonymousfalse
\newif\iffullversion \fullversiontrue
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage{color}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bcprules, proof}
\usepackage{fancybox}
\usepackage{mathtools}
\usepackage{float}
\usepackage{xparse}
\usepackage{lscape}
\usepackage{xspace}
\usepackage{url}

@@

\begin{document}
%
\title{A Dependently Typed Multi-Stage Calculus%
	\ifnotanonymous\thanks{Supported by organization x.}\fi
}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\ifnotanonymous
	\author{Akira Kawata\inst{1} \and
		Atsushi Igarashi\inst{2}\orcidID{0000-0002-5143-9764}}
	% %
	\authorrunning{A. Kawata, A. Igarashi}
	% % First names are abbreviated in the running head.
	% % If there are more than two authors, 'et al.' is used.
	% %
	\institute{Graduate School of Informatics, Kyoto University, Kyoto, Japan\\
		\email{akira@fos.kuis.kyoto-u.ac.jp} \and
		\email{igarashi@kuis.kyoto-u.ac.jp}
	}
\fi
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}

	We study a dependently typed extension of a multi-stage programming
	language \`a la MetaOCaml, which supports quasi-quotation and
	cross-stage persistence for manipulation of code fragments as
	first-class values and eval for execution of programs dynamically
	generated by the code manipulation.  Dependent types are expected to
	bring to multi-stage programming enforcement of strong
	invariants---beyond simple type safety---on the behavior of
	dynamically generated code.  An extension is, however, not trivial
	because a type system would have to take stages---roughly speaking,
	the number of surrounding quotations---of types into account.

@@

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\newif\ifnotanonymous \notanonymoustrue
\newif\iffullversion \fullversiontrue
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage{color}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bcprules, proof}
\usepackage{fancybox}
\usepackage{mathtools}
\usepackage{float}
\usepackage{xparse}
\usepackage{lscape}
\usepackage{xspace}
\usepackage{url}
\usepackage{bcpproof}

@@

\begin{document}
%
\title{A Dependently Typed Multi-Stage Calculus%
	% \ifnotanonymous\thanks{Supported by organization x.}\fi
}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\ifnotanonymous
	\author{Akira Kawata\inst{1} \and
		Atsushi Igarashi\inst{1}\orcidID{0000-0002-5143-9764}}
	% %
	\authorrunning{A. Kawata, A. Igarashi}
	% % First names are abbreviated in the running head.
	% % If there are more than two authors, 'et al.' is used.
	% %
	\institute{Graduate School of Informatics, Kyoto University, Kyoto, Japan\\
          \email{akira@fos.kuis.kyoto-u.ac.jp} \\
          \email{igarashi@kuis.kyoto-u.ac.jp}
	}
\fi
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}

	We study a dependently typed extension of a multi-stage programming language
\`a la MetaOCaml, which supports quasi-quotation and cross-stage persistence
for manipulation of code fragments as first-class values and an evaluation
construct for execution of programs dynamically generated by this code
manipulation. Dependent types are expected to bring to multi-stage
programming enforcement of strong invariants---beyond simple type safety---on
the behavior of dynamically generated code. An extension is, however, not
trivial because a type system would have to take stages---roughly speaking,
the number of surrounding quotations---of types into account.

@@

\input{text/acknowledgments.tex}

### 2019-08-14 text/formal-defnition.tex
Found a bug in the defnition of Redex.
Reflected comments on typing from reviwer 1,

A kind, which is used to classify types, is either $*$, the kind of
proper types (types that terms inhabit), or $\Pi x\colon\tau.K$, the kind
of type operators that takes $x$ as an argument of type $\tau$ and returns a type
of kind $K$.
% Types
% of terms have kind $*$ and dependent types have $\Pi$-kinds.  For
% example, $\lambda x:\I.x$ has type $\Pi x:\I.\I$, which has $*$ kind.
% \red{この段落は短いので型の段落と結合するか?}
% Types
A type is a type-level constant $X$, which is declared in the signature with its kind, a dependent function type $\Pi x:\tau.\sigma$,
an application $\tau\ M$ of a type (operator of $\Pi$-kind) to a term, a code type $\TW_\alpha \tau$, or an $\alpha$-closed type $\F\alpha.\tau$.
An example of an application of a type (operator) of $\Pi$-kind to a term is $\text{Vector}\ 10$; it is well kinded
if, say, type-level constant $\text{Vector}$ has kind $\Pi x:\I.*$.
A code type $\TW_\alpha \tau$ is for a code fragment of a term of type $\tau$.
An $\alpha$-closed type, when used with $\TW_\alpha$, represents runnable code.

@@

We adopt the tradition of \LLF-like systems, where types of constants and kinds of type-level constants are globally declared in a signature $\Sigma$, which
is a sequence of declarations of the form $c:\tau$ and $X::K$.
For example, when we use Boolean in \LMD, $\Sigma$ includes $\B :: *, \textrm{true}:\B, \textrm{false}:\B$.
Type environments are sequences of triples of a variable, its type, and its stage.
We write \(\textit{dom}(\Sigma)\) and \(\textit{dom}(\Gamma)\)
for the set of (type-level) constants and variables declared in \(\Sigma\) and \(\Gamma\), respectively.   As in other multi-stage calculi~\cite{taha2003environment,Tsukada,Hanada2014},
a variable declaration is associated with a stage
so that a variable can be referenced only at the declared stage.
On the contrary, constants and type-level constants are \emph{not} associated with stages;
so, they can appear at any stage.
We define well-formed signature and well-formed type environments later.

@@

The relation $\longrightarrow_\beta$ represents ordinary $\beta$-reduction in the \(\lambda\)-calculus; the relation
$\longrightarrow_\blacklozenge$ represents that quotation $\TB_\alpha M$ is canceled by escape and $M$ is spliced into the code fragment surrounding the escape;
the relation $\longrightarrow_\Lambda$ means that a stage abstraction applied to  stage $A$ reduces to the body of the abstraction
where $A$ is substituted for the stage variable.
There is no reduction rule for CSP as with Hanada and Igarashi \cite{Hanada2014}.
The CSP operator $\%_\alpha$ disappears when $\varepsilon$ is substituted for $\alpha$.
We show an example of a reduction sequence below.
Underlines show the redexes.
\begin{align*}
     & \hspace{10mm} \underline{(\lambda i:\I\to\I.(\Lambda\alpha.\TB_\alpha (\%_\alpha i\ 1 + (\TBL_\alpha \TB_\alpha 3))\ \varepsilon))\ (\lambda x:\I.x)} \\
     & \longrightarrow_\blacklozenge (\Lambda\alpha.\TB_\alpha (\%_\alpha (\lambda x:\I.x)\ 1 + (\underline{\TBL_\alpha \TB_\alpha 3})))\ \varepsilon        \\
     & \longrightarrow_\beta \underline{(\Lambda\alpha.\TB_\alpha (\%_\alpha (\lambda x:\I.x)\ 1 + 3))\ \varepsilon}                                         \\
     & \longrightarrow_\Lambda \underline{(\lambda x:\I.x)\ 1} + 3                                                                                           \\
     & \longrightarrow_\beta 1 + 3                                                                                                                           \\
     & \longrightarrow^* 4
\end{align*}

@@

Typing rules of \LMD are shown in Figure~\ref{fig:typing-rules}.
The rule \TConst{} means that a constant can appear at any stage.
% For example, if we have a signature $\Sigma$ which is
% $\textrm{bool} :: *, \textrm{true}: \textrm{bool}, \textrm{false}:
%     \textrm{bool}$, the derivation tree in
% Figure~\ref{fig:tconst-derivation-tree} is admissible.
The rules \TVar,\TAbs, and \TApp{} are almost the same as those in the simply typed
lambda calculus or \LLF.  Additional conditions are that subterms must be
typed at the same stage (\TAbs{} and \TApp); the type
annotation/declaration on a variable has to be a proper type of kind
$*$ (\TAbs) at the stage where it is declared (\TVar{} and \TAbs).
% \AI{Don't we need signature well formedness?  If $c:\tau \in \Sigma$, then $\tau$ should be a well-formed type under ... what?}
% \AI{... or \(\lambda\)LF?  We are working in a dependently type system...}

@@

Since the syntax of kinds, types, and terms is mutually recursive,
corresponding notions of equivalence are also mutually recursive.
They are congruences closed under a few axioms for term equivalence.
So, the rules for kind and type equivalences are not very interesting
except implicit CSP is allowed.
We show a few representative rules below.

@@

Interestingly, Hanada and Igarashi~\cite{Hanada2014} rejected the idea of
reduction that removes $\%_\alpha$ when they developed \LTP{}, as such
reduction does not match the operational behavior of the CSP operator
in implementation.  However, as an equational system for multi-stage
programs, the rule \QPercent makes sense.

@@

\begin{definition}[Redex]
  The sets of $\varepsilon$-redexes (ranged over by $R^\varepsilon$) and $\alpha$-redexes (ranged over by $R^\alpha$) are defined by the following grammar.
  \begin{align*}
     & R^\varepsilon ::= (\lambda x:\tau.M)\ v^\varepsilon \mid (\Lambda\alpha.v^\varepsilon)\ \varepsilon \\
     & R^\alpha      ::=\ \TBL_\alpha \TB_\alpha M                                                         \\
  \end{align*}
\end{definition}

\begin{definition}[Evaluation Context]
  Let $B$ be either \(\varepsilon\) or a stage variable \(\beta\).
  The family of sets $ECtx^A_B$ of evaluation contexts, ranged over by $E^A_B$, is defined by the following grammar (in which $A'$ stands for a non-empty stage).
  %  $A$ is assumed to be nonempty (but $B,A'$ can be empty).
  \begin{align*}
    E^\varepsilon_B \in ECtx^\varepsilon_B & ::= \square\ (\text{if\ } B = \varepsilon)
    \mid E^\varepsilon_B\ M \mid v^\varepsilon\ E^\varepsilon_B \mid \TB_\alpha E^\alpha_B
    \mid \Lambda\alpha.E^\varepsilon_B \mid E^\varepsilon_B\ A                                                                                    \\
    E^{A'}_B \in ECtx^{A'}_B               & ::= \square\ (\text{if } A' = B) \mid \lambda x:\tau.E^{A'}_B \mid E^{A'}_B\ M \mid v^{A'}\ E^{A'}_B \\
                                           & \mid \TB_\alpha E^{A'\alpha}_B \mid \TBL_\alpha E^{A}_B \ (\text{where } A\alpha = A')               \\
                                           & \mid \Lambda\alpha.E^{A'}_B \mid E^{A'}_B\ A \mid \%_\alpha\ E^{A'}_B \ (\text{where } A\alpha = A')
  \end{align*}
\end{definition}

>>>>

A kind, which is used to classify types, is either $*$, the kind of
proper types (types that terms inhabit), or $\Pi x\colon\tau.K$, the kind
of type operators that takes $x$ as an argument of type $\tau$ and returns a type
of kind $K$.
% Types
% of terms have kind $*$ and dependent types have $\Pi$-kinds.  For
% example, $\lambda x:\I.x$ has type $\Pi x:\I.\I$, which has $*$ kind.
% \red{この段落は短いので型の段落と結合するか?}
% Types
A type is a type-level constant $X$, which is declared in the signature with its kind, a dependent function type $\Pi x:\tau.\sigma$,
an application $\tau\ M$ of a type (operator of $\Pi$-kind) to a term, a code type $\TW_\alpha \tau$, or an $\alpha$-closed type $\F\alpha.\tau$.
An example of an application of a type (operator) of $\Pi$-kind to a term is $\text{Vector}\ 10$; it is well kinded
if, say, the type-level constant $\text{Vector}$ has kind $\Pi x:\I.*$.
A code type $\TW_\alpha \tau$ is for a code fragment of a term of type $\tau$.
An $\alpha$-closed type, when used with $\TW_\alpha$, represents runnable code.

@@

We adopt the tradition of \LLF-like systems, where types of constants and
kinds of type-level constants are globally declared in a signature $\Sigma$,
which is a sequence of declarations of the form $c:\tau$ and $X::K$. For
example, when we use Boolean in \LMD, $\Sigma$ includes $\B :: *,
\textrm{true}:\B, \textrm{false}:\B$. Type environments are sequences of
triples of a variable, its type, and its stage. We write
\(\textit{dom}(\Sigma)\) and \(\textit{dom}(\Gamma)\) for the set of
(type-level) constants and variables declared in \(\Sigma\) and \(\Gamma\),
respectively. As in other multi-stage
calculi~\cite{taha2003environment,Tsukada,Hanada2014}, a variable declaration
is associated with a stage so that a variable can be referenced only at the
declared stage. On the contrary, constants and type-level constants are
\emph{not} associated with stages; so, they can appear at any stage. We
define well-formed signatures and well-formed type environments later.

@@

The relation $\longrightarrow_\beta$ represents ordinary $\beta$-reduction in the \(\lambda\)-calculus; the relation
$\longrightarrow_\blacklozenge$ represents that quotation $\TB_\alpha M$ is canceled by escape and $M$ is spliced into the code fragment surrounding the escape;
the relation $\longrightarrow_\Lambda$ means that a stage abstraction applied to  stage $A$ reduces to the body of the abstraction
where $A$ is substituted for the stage variable.
There is no reduction rule for CSP as with Hanada and Igarashi \cite{Hanada2014}.
The CSP operator $\%_\alpha$ disappears when $\varepsilon$ is substituted for $\alpha$.
We show an example of a reduction sequence below.
Underlines show the redexes.
\begin{align*}
     & \hspace{10mm} \underline{(\lambda i:\I\to\I.(\Lambda\alpha.\TB_\alpha (\%_\alpha i\ 1 + (\TBL_\alpha \TB_\alpha 3))\ \varepsilon))\ (\lambda x:\I.x)} \\
     & \longrightarrow_\beta (\Lambda\alpha.\TB_\alpha (\%_\alpha (\lambda x:\I.x)\ 1 + (\underline{\TBL_\alpha \TB_\alpha 3})))\ \varepsilon        \\
     & \longrightarrow_\blacklozenge \underline{(\Lambda\alpha.\TB_\alpha (\%_\alpha (\lambda x:\I.x)\ 1 + 3))\ \varepsilon}                                         \\
     & \longrightarrow_\Lambda \underline{(\lambda x:\I.x)\ 1} + 3                                                                                           \\
     & \longrightarrow_\beta 1 + 3                                                                                                                           \\
     & \longrightarrow^* 4
\end{align*}

@@

The typing rules of \LMD are shown in Figure~\ref{fig:typing-rules}.
The rule \TConst{} means that a constant can appear at any stage.
% For example, if we have a signature $\Sigma$ which is
% $\textrm{bool} :: *, \textrm{true}: \textrm{bool}, \textrm{false}:
%     \textrm{bool}$, the derivation tree in
% Figure~\ref{fig:tconst-derivation-tree} is admissible.
The rules \TVar,\TAbs, and \TApp{} are almost the same as those in the simply typed
lambda calculus or \LLF.  Additional conditions are that subterms must be
typed at the same stage (\TAbs{} and \TApp); the type
annotation/declaration on a variable has to be a proper type of kind
$*$ (\TAbs) at the stage where it is declared (\TVar{} and \TAbs).
% \AI{Don't we need signature well formedness?  If $c:\tau \in \Sigma$, then $\tau$ should be a well-formed type under ... what?}
% \AI{... or \(\lambda\)LF?  We are working in a dependently type system...}

@@

Since the syntax of kinds, types, and terms is mutually recursive,
the corresponding notions of equivalence are also mutually recursive.
They are congruences closed under a few axioms for term equivalence.
So, the rules for kind and type equivalences are not very interesting, 
except that implicit CSP is allowed.
We show a few representative rules below.

@@

Interestingly, Hanada and Igarashi~\cite{Hanada2014} rejected the idea of
reduction that removes $\%_\alpha$ when they developed \LTP{}, as such
reduction does not match the operational behavior of the CSP operator
in implementations.  However, as an equational system for multi-stage
programs, the rule \QPercent makes sense.

@@

\begin{definition}[Redex]
  The sets of $\varepsilon$-redexes (ranged over by $R^\varepsilon$) and $\alpha$-redexes (ranged over by $R^\alpha$) are defined by the following grammar.
  \begin{align*}
     & R^\varepsilon ::= (\lambda x:\tau.M)\ v^\varepsilon \mid (\Lambda\alpha.v^\varepsilon)\ \varepsilon \\
     & R^\alpha      ::=\ \TBL_\alpha \TB_\alpha v^\alpha                                                         \\
  \end{align*}
\end{definition}

\begin{definition}[Evaluation Context]
  Let $B$ be either \(\varepsilon\) or a stage variable \(\beta\).
  The family of sets $ECtx^A_B$ of evaluation contexts, ranged over by $E^A_B$, is defined by the following grammar (in which $A'$ stands for a non-empty stage).
  %  $A$ is assumed to be nonempty (but $B,A'$ can be empty).
  \begin{align*}
    E^\varepsilon_B \in ECtx^\varepsilon_B & ::= \square\ (\text{if\ } B = \varepsilon)
    \mid E^\varepsilon_B\ M \mid v^\varepsilon\ E^\varepsilon_B \mid \TB_\alpha E^\alpha_B
    \mid \Lambda\alpha.E^\varepsilon_B \mid E^\varepsilon_B\ A                                                                                    \\
    E^{A'}_B \in ECtx^{A'}_B               & ::= \square\ (\text{if } A' = B) \mid \lambda x:\tau.E^{A'}_B \mid E^{A'}_B\ M \mid v^{A'}\ E^{A'}_B \\
                                           & \mid \TB_\alpha E^{A'\alpha}_B \mid \TBL_\alpha E^{A}_B \ (\text{where } A\alpha = A')               \\
                                           & \mid \Lambda\alpha.E^{A'}_B \mid E^{A'}_B\ A \mid \%_\alpha\ E^{A}_B \ (\text{where } A\alpha = A')
  \end{align*}
\end{definition}

### 2019-08-14 text/informal-overview.tex
Reflected comments on typing from reviwer 1,

The subscript $\alpha$ in $\TB_\alpha$ and $\TBL_\alpha$ is a \textit{stage
  variable}\footnote{%
  In Hanada and Igarashi~\cite{Hanada2014}, it was called a
  \textit{transition variable}, which is derived from correspondence
  to modal logic, studied by Tsukada and Igarashi~\cite{Tsukada}.} and
a sequence of stage variables is called a \textit{stage}.  Intuitively,
they represent the depth of nested brackets.  Stage variables can be
abstracted by $\Lambda\alpha.M$ and instantiated by an application
$M\ A$ to stages.  For example,
$\Lambda\alpha.\TB_\alpha ((\lambda x:\I.x+10)\ 5)$ is a code value,
where \(\alpha\) is abstracted.  If it is applied to \(\alpha_1 \cdots \alpha_n\), \(\TB_\alpha\) becomes \(\TB_{\alpha_1} \cdots \TB_{\alpha_n}\); in particular,
if \(n = 0\), \(\TB_\alpha\) disappears.  So, an
application of $\Lambda\alpha.\TB_\alpha ((\lambda x:\I.x+10)\ 5)$
to the empty sequence \(\varepsilon\) reduces to
(unquoted) \((\lambda x:\I.x+10)\ 5\) and to 15.  In other words, application of of
\(\Lambda\)-abstraction to $\varepsilon$ corresponds to \texttt{run}.
% If \(\alpha\) is instantiated with, say, \(\beta\gamma\), then
% \(\TB_\alpha\) becomes nested brackets \(\TB_{\beta} \TB_\gamma\).
This is expressed by the following reduction rule:
\begin{align*}
	(\Lambda\alpha.M)\ A \longrightarrow M[\alpha\mapsto A]
\end{align*}
where stage substitution \([\alpha \mapsto A]\) manipulates nesting of
\(\TB_\alpha\) and \(\TBL_\alpha\) (and also \(\%_\alpha\) as we see later).

@@

The type system of \LTP uses code types---the type of code of type
\(\tau\) is written \(\TW_\alpha \tau\)---for typing \(\TB_\alpha\),
\(\TBL_\alpha\) and \(\%_\alpha\).  It takes stages into account: A
variable declaration (written $x:\tau@A$) in a type environment is associated with its
declared stage $A$ as well as its type $\tau$ and the type judgement of \LTP is of
the form $\G \vdash M : \tau@A$, in which $A$ stands for the stage
of term $M$.\footnote{%
  In Hanada and Igarashi~\cite{Hanada2014}, it is written
  $\G \vdash^A M : \tau$.
  }
For example,
$y:\I@\alpha \vdash (\lambda x:\I.y) : \I \to \I @ \alpha$ holds, but
$y:\I@\alpha \vdash (\lambda x:\I.y) : \I \to \I @ \varepsilon$ does
not because the latter uses $y$ at stage \(\varepsilon\) but $y$ is
declared at $\alpha$.  Quotation \(\TB_\alpha M\) is given type
\(\TW_\alpha \tau\) at stage \(A\) if \(M\) is given type \(\tau\) at
stage \(A\alpha\); unquote \(\TBL_\alpha M\) is given type \(\tau\)
at stage \(A\alpha\) if \(M\) is given type \(\TW_\alpha \tau\) at
stage \(A\alpha\); and CSP \(\%_\alpha M\) is give type \(\tau\)
at stage \(A\alpha\) if \(M\) is given type \(\tau\) at \(A\).
These are expressed by the following typing rules.
\begin{center}
	\infrule{\G\vdash M:\tau @{A\alpha}}{\G\vdash \TB_{\alpha}M:\TW_{\alpha}\tau @A} \hfil
	\infrule{\G\vdash M:\TW_{\alpha}\tau @A}{\G\vdash \TBL_{\alpha}M:\tau @{A\alpha}} \hfil
	\infrule{\G\vdash M: \tau @A}{\G\vdash \%_{\alpha}M:\tau @{A\alpha}}
\end{center}
% \TTB, corresponding to brackets, means 
% if $M$ is typed $\tau$ at stage $A\alpha$ then $\TB_{\alpha}M$, quoted $M$, is typed $\TW_{\alpha}\tau$ at $A$.
% \TTBL\ is converse of \TTB.

@@

Second, an indexed type such as Vector\ 3 or Vector\ $x$ is well
formed only at the stage(s) where the index term is well typed.  Since
constant \(3\) is well typed at every stage (if it is declared in the
signature), Vector\ 3 is a well-formed type at every stage, too.
However, Vector\ $M$ is well formed only at the stage where index term
$M$ is typed.  Thus, the kinding judgment
of \LMD takes the form \(\G\V \tau :: K @ A\), where stage $A$ stands for
where \(\tau\) is well formed.  For example,
given \(\text{Vector}:: \I \rightarrow *\) in the signature \(\Sigma\),
\(x:\I@\varepsilon \V \text{Vector }x :: * @\varepsilon\) can be
derived but
neither
\(x:\I@\alpha \V \text{Vector }x :: * @\varepsilon\)
nor 
\(x:\I@\epsilon \V \text{Vector }x :: * @\alpha\)
can be.

At first, the restriction above sounds too severe, because a term like
\(\lambda x:\I. \TB_\alpha (\lambda y:\text{Vector }x.M) \), which
models a typical code generator which takes the size $x$ and returns
code for vector manipulation specialized to the given size.  It seems
crucial for \(y\) to be given a type indexed by $x$.  We can address
this problem by CSP---In fact,
$\text{Vector }x$ is not well formed at $\alpha$ under $x:\I@\epsilon$
but $\text{Vector }(\%_\alpha x)$ is!  So, we can still write
\(\lambda x:\I. \TB_\alpha (\lambda y:\text{Vector }(\%_\alpha x).M) \)
for the typical sort of code generators.

>>>>

The subscript $\alpha$ in $\TB_\alpha$ and $\TBL_\alpha$ is a \textit{stage
  variable}\footnote{%
  In Hanada and Igarashi~\cite{Hanada2014}, it was called a
  \textit{transition variable}, which is derived from correspondence
  to modal logic, studied by Tsukada and Igarashi~\cite{Tsukada}.} and
a sequence of stage variables is called a \textit{stage}.  Intuitively,
they represent the depth of nested brackets.  Stage variables can be
abstracted by $\Lambda\alpha.M$ and instantiated by an application
$M\ A$ to stages.  For example,
$\Lambda\alpha.\TB_\alpha ((\lambda x:\I.x+10)\ 5)$ is a code value,
where \(\alpha\) is abstracted.  If it is applied to \(\alpha_1 \cdots \alpha_n\), \(\TB_\alpha\) becomes \(\TB_{\alpha_1} \cdots \TB_{\alpha_n}\); in particular,
if \(n = 0\), \(\TB_\alpha\) disappears.  So, an
application of $\Lambda\alpha.\TB_\alpha ((\lambda x:\I.x+10)\ 5)$
to the empty sequence \(\varepsilon\) reduces to
(unquoted) \((\lambda x:\I.x+10)\ 5\) and to 15.  In other words, application of of
\(\Lambda\)-abstraction to $\varepsilon$ corresponds to \texttt{run}.
% If \(\alpha\) is instantiated with, say, \(\beta\gamma\), then
% \(\TB_\alpha\) becomes nested brackets \(\TB_{\beta} \TB_\gamma\).
This is expressed by the following reduction rule:
\begin{align*}
	(\Lambda\alpha.M)\ A \longrightarrow M[\alpha\mapsto A]
\end{align*}
where stage substitution \([\alpha \mapsto A]\) manipulates the nesting of
\(\TB_\alpha\) and \(\TBL_\alpha\) (and also \(\%_\alpha\) as we see later).

@@

The type system of \LTP uses code types---the type of code of type
\(\tau\) is written \(\TW_\alpha \tau\)---for typing \(\TB_\alpha\),
\(\TBL_\alpha\) and \(\%_\alpha\).  It takes stages into account: a
variable declaration (written $x:\tau@A$) in a type environment is associated with its
declared stage $A$ as well as its type $\tau$ and the type judgement of \LTP is of
the form $\G \vdash M : \tau@A$, in which $A$ stands for the stage
of term $M$.\footnote{%
  In Hanada and Igarashi~\cite{Hanada2014}, it is written
  $\G \vdash^A M : \tau$.
  }
For example,
$y:\I@\alpha \vdash (\lambda x:\I.y) : \I \to \I @ \alpha$ holds, but
$y:\I@\alpha \vdash (\lambda x:\I.y) : \I \to \I @ \varepsilon$ does
not because the latter uses $y$ at stage \(\varepsilon\) but $y$ is
declared at $\alpha$.  Quotation \(\TB_\alpha M\) is given type
\(\TW_\alpha \tau\) at stage \(A\) if \(M\) is given type \(\tau\) at
stage \(A\alpha\); unquote \(\TBL_\alpha M\) is given type \(\tau\)
at stage \(A\alpha\) if \(M\) is given type \(\TW_\alpha \tau\) at
stage \(A\alpha\); and CSP \(\%_\alpha M\) is give type \(\tau\)
at stage \(A\alpha\) if \(M\) is given type \(\tau\) at \(A\).
These are expressed by the following typing rules.
\begin{center}
	\infrule{\G\vdash M:\tau @{A\alpha}}{\G\vdash \TB_{\alpha}M:\TW_{\alpha}\tau @A} \hfil
	\infrule{\G\vdash M:\TW_{\alpha}\tau @A}{\G\vdash \TBL_{\alpha}M:\tau @{A\alpha}} \hfil
	\infrule{\G\vdash M: \tau @A}{\G\vdash \%_{\alpha}M:\tau @{A\alpha}}
\end{center}
% \TTB, corresponding to brackets, means 
% if $M$ is typed $\tau$ at stage $A\alpha$ then $\TB_{\alpha}M$, quoted $M$, is typed $\TW_{\alpha}\tau$ at $A$.
% \TTBL\ is converse of \TTB.

@@

Second, an indexed type such as Vector\ 3 or Vector\ $x$ is well
formed only at the stage(s) where the index term is well typed.  Since
constant \(3\) is well typed at every stage (if it is declared in the
signature), Vector\ 3 is a well-formed type at every stage, too.
However, Vector\ $M$ is well formed only at the stage where index term
$M$ is typed.  Thus, the kinding judgment
of \LMD takes the form \(\G\V \tau :: K @ A\), where stage $A$ stands for
where \(\tau\) is well formed.  For example,
given \(\text{Vector}:: \I \rightarrow *\) in the signature \(\Sigma\),
\(x:\I@\varepsilon \V \text{Vector }x :: * @\varepsilon\) can be
derived but
neither
\(x:\I@\alpha \V \text{Vector }x :: * @\varepsilon\)
nor 
\(x:\I@\varepsilon \V \text{Vector }x :: * @\alpha\)
can be.

At first, the restriction above sounds too severe, because a term like
\(\lambda x:\I. \TB_\alpha (\lambda y:\text{Vector }x.M) \), which
models a typical code generator which takes the size $x$ and returns
code for vector manipulation specialized to the given size.  It seems
crucial for \(y\) to be given a type indexed by $x$.  We can address
this problem by CSP---In fact,
$\text{Vector }x$ is not well formed at $\alpha$ under $x:\I@\varepsilon$
but $\text{Vector }(\%_\alpha x)$ is!  So, we can still write
\(\lambda x:\I. \TB_\alpha (\lambda y:\text{Vector }(\%_\alpha x).M) \)
for the typical sort of code generators.
